---
title: "Spanish beaches"
author: "Sergio Aguado"
highlighter: highlight.js
output: html_document
job: null
knit: slidify::knit2slides
mode: selfcontained
hitheme: tomorrow
subtitle: One of the most important resources for attracting tourism
framework: io2012
widgets: shiny
---

## Introduction

In 2014, Spain was the second classified in the ranking of ["International tourism receipts"](http://dtxtq4w60xqpw.cloudfront.net/sites/all/files/pdf/unwto_barom15_02_april_excerpt__0.pdf). Culture, history, nature, sun are some of the things, that attract tourists from all over the world. But without any doubt, the Spanish beaches are the most important natural resource for the tourism sector.

This is study tries to describe the Spanish beaches: Length, width, vegetation, 
urbanization and occupation level, and other kind of equipments.

The data set used for this research is "Guía de playas" from the open data initiative from
the Spanish government. The data set has been translated in English because is the language of
this course. 

<br>
You can download the data set from [here](http://datos.gob.es/catalogo/guia-de-playas)

--- .class #id 
## Summary

Spain has 3.463 beaches shared in 12 Autonomous Communities with a total length of
1871.6 kilometers. But to make a better idea about the Spanis beaches and their 
equipments, take a look at the following table

```{r echo=FALSE}
library(plyr)
suppressPackageStartupMessages(library(googleVis))
beaches <- read.csv("data/beaches.csv", header=T, sep=",")

length <- round(sum(beaches$Length, na.rm = T) / 1000,1)


```

```{r echo=F, results = 'asis'}
num.beach <- as.numeric(dim(beaches)[1])
# Blue flag beaches
blueFlag <- sum(beaches$Blue_flag ==T, na.rm = T)
blueFlag <- c("Blue flag*", blueFlag, 
              paste(round((blueFlag/num.beach)*100,0)," %", sep=""))

# Beaches with SOS
sos <- sum(beaches$SOS == T, na.rm=T)
sos <- c("SOS", sos, paste(round((sos/num.beach)*100,0)," %", sep=""))

# Disabled access
disabled <- sum(beaches$Disabled_access == T, na.rm=T)
disabled <- c("Disabled Access", disabled, 
              paste(round((disabled/num.beach)*100,0)," %", sep=""))

# Toilets
toilets <- sum(beaches$Toilets == T, na.rm=T)
toilets <- c("Toilets", toilets, 
             paste(round((toilets/num.beach)*100,0)," %", sep=""))

# Showers
showers <- sum(beaches$Showers == T, na.rm=T)
showers <- c("Showers", showers, 
             paste(round((showers/num.beach)*100,0)," %", sep=""))

# Tourist offices
touristOffice <- sum(beaches$Tourism_office == T, na.rm=T)
touristOffice <- c("Tourist Offices", touristOffice, 
                   paste(round((touristOffice/num.beach)*100,0)," %", sep=""))

# Children areas
childrenArea <- sum(beaches$Children_area == T, na.rm=T)
childrenArea <- c("Children areas", childrenArea, 
                   paste(round((childrenArea/num.beach)*100,0)," %", sep=""))

# Surfing
surf <- sum(beaches$Surf == T, na.rm=T)
surf <- c("Surf", surf, 
                   paste(round((surf/num.beach)*100,0)," %", sep=""))


# Create data table
Btable <- as.data.frame(rbind(blueFlag, sos, disabled, toilets, showers, 
                              touristOffice, childrenArea, surf))

names(Btable) <- c("Services","Number of Beaches", "% of total")

print(gvisTable(Btable, 
                     options=list(page='enable',
                                  width="80%")), tag = 'chart')
``` 


<br>
*Spain is the top 1 in blue flag award. [Ranking](http://www.blueflag.org/menu/awarded-sites)

--- .class #id 

## Average width and length of the beaches

```{r echo=FALSE}

library(plyr)
beaches <- read.csv("data/beaches.csv", header=T, sep=",")
# Aggregate the legth and width beaches data by Autonomous_region
beachesByRegion <- aggregate(beaches[,c(10,11)], by=list(Region=beaches$Autonomous_region), 
                             FUN = mean, na.action=na.pass, na.rm=TRUE)
num.beaches <- count(beaches,vars="Autonomous_region")
names(num.beaches) <- c("Region", "num.beaches")
beachesByRegion <- merge(beachesByRegion, num.beaches, by = "Region")
beachesByRegion$Length <- round(beachesByRegion$Length,0)
beachesByRegion$Width <- round(beachesByRegion$Width, 0)
names(beachesByRegion) <- c("Region", "Avg.Length", "Avg.Width", "num.beaches")


```

```{r echo=F, results = 'asis'}
listOptions <- list(height=500,
                    width=1000,
                    hAxis="{title:'Avg beaches width (m)'}",
                    vAxis="{title:'Avg beaches length (m)'}")
                    #chartArea="{width:80%}")

print(gvisBubbleChart(beachesByRegion, idvar="Region", 
                          yvar="Avg.Length", xvar="Avg.Width",
                          colorvar="Region", sizevar="num.beaches", 
                          options = listOptions),
      tag = 'chart')
```

<br>
The data is shown by Autonomous Communities. 


--- .class #id 
## Shiny application

The following shiny application is only the first step of this research and the goal 
is to be available of discovering beaches through their equiptments and properties. 
You will only see the filters for:
- Occupation level
- Urbanization level
- Length
- Width
- Vegetation

There are only these filters in the application due to the delivery date of the 
course Developing data product from the [Data Science Specialization](https://www.coursera.org/specialization/jhudatascience/1?utm_medium=catalog)


--- .class #id 
## How to use the shiny application

You will see on the left of the web page the filters for filtering the data. By
default is selected "Low" in Occupation level and "Semiurban" in Urbanization level.
It is because the google map doesn't show the points if there are too many. 

After filtering, you can click on the map's points and see more information about
the beaches. 

<br>
Shiny application: [Beaches Project](https://sergioaguado.shinyapps.io/Beaches_project/)

<br>
I hope that you enjoy this application by discovering nice beaches for going this
summer. 


